---
title: "Data simulation and example analysis for _Scene Visualization: Transfer from Text to Visuals_"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup}
library(lme4)
library(lattice)
set.seed(1614)
```

# Description

This is a simulated data set with accompanying data analysis which was used as
basis for a power simulation for an experiment at the Leibniz-Institut für
Wissensmedien (IWM).

Underlying idea of the study: People construct mental representations of a scene
while reading, meaning textual content is, e.g., translated into visual mental
representations. This is tested by switching the presentation modality between
the learning and the test phase. During the learning phase, subjects are
presented with a scene description in text form. In the test phase, the scenes
are presented either in text or in graphical form.

Transfer of text-based modality to visual modality:

* Are participants able to identify information in a visual representation of a
  scene they read? 

* How accurate is it?

Participants read 25 sentences that can be either true (15) or false (10).
Afterwards, they have to judge if a scene is true or false. This scene is either
presented as text or as a picture.

The dependent variable is correct answer: true sentence is classified as true.

The manipulated conditions are:

* representation: text vs. picture

* veracity of sentence: true vs. false (indicated by color of the sentence)


# Hypothesis

Participants will be better in identifying _false_ information in a text-based
representation than participants who must identify false information in a
graphical representation of the scene. No difference for true information is
expected.

```{r hypothesis, fig.align = "center", fig.height = 5, fig.width = 5}
# Hypothesis: 0.1 probability difference for true vs. false for graphic representation
design <- data.frame(representation = factor(rep(c("text", "graphic"), each = 2),
                                             levels = c("text", "graphic")),
                     veracity = factor(rep(c("true", "false"), 2),
                                       levels = c("true", "false")),
                     prob = c(0.55, 0.55, 0.50, 0.40))

xyplot(prob ~ representation, data = design, groups = veracity,
       type = "b", ylim = 0:1, auto.key = list(space = "top"))
knitr::kable(design)
```

# Data simulation

```{r datasim}
# Fixed effects
X <- model.matrix( ~ representation * veracity, data = design)
(fix <- as.vector(solve(X) %*% qlogis(design$prob)))

# Random effects (taken from previous study)
ran <- c("id.(Intercept)" = 0.5, "item.(Intercept)" = 0.3)

# Simulate data
n <- 150

dat <- data.frame(
    id              = factor(rep(1:n, each = 25)),
    item            = factor(paste(rep(1:5, each = 25), 1:25, sep = ":")),
    representation  = factor(rep(c("text", "graphic"), each = 25)),
    veracity        = factor(rep(c("true", "false"), c(15, 10)))
)

dat$representation <- relevel(dat$representation, ref = "text")
dat$veracity <- relevel(dat$veracity, ref = "true")

dat$resp <- simulate( ~ representation * veracity + (1|id) + (1|item),
                     newdata   = dat,
                     newparams = list(beta = fix, theta = ran),
                     family    = binomial)$sim_1
```

## Descriptives

```{r descriptives, fig.align = "center", fig.height = 5, fig.width = 5}
# Frequency table of observations per condition
(tab1 <- xtabs( ~ representation + veracity, dat))
# Frequency table of correct answers per condition
(tab2 <- xtabs(resp ~ representation + veracity, dat))

datm <- as.data.frame(tab2 / tab1)

xyplot(Freq ~ representation, data = datm, groups = veracity,
       type = "b", ylim = 0:1, auto.key = list(space = "top"))

knitr::kable(datm)
```

# Analysis with `lme4::glmer()`

```{r analysis}

m1 <- glmer(resp ~ representation * veracity + (1|id) + (1|item), dat,
            family = binomial)

summary(m1)
```

