---
title: "Data simulation and example analysis for _Event Segmentation with Dynamic Maps: Influence of Framing and Spatial Pattern of Change_"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup}
set.seed(1331)

library(dplyr)
library(lme4)
```

# Description

This is a simulated data set with accompanying data analysis which was used as
basis for a power simulation for an experiment at the Leibniz-Institut für
Wissensmedien (IWM).

The experiment studies segmentation behavior with dynamic maps depicting the
development of insect populations on fictional islands. Specifically, how the
framing of the insect species as invasive or endangered, the trend of the shown
population development and the spatial pattern of the temporal change impact
segmentation agreement is investigated.

## Dependent variable

Participants are tasked to view animated maps and press the space bar key
whenever they perceive that a meaningful unit has ended and a new one has begun.
Therefore, the dependent variable is the timing of these key presses in
milliseconds relative to the stimulus onset. From there, we calculate
segmentation agreement as follows:

1. For each stimulus and participant, convert the timing of key presses into 30
   1-second bins of 0s and 1s, indicating whether the participant has pressed
   the key during that second

2. Calculate the group norm for that stimulus by calculating the proportion of
   all other participants that have pressed the key in each 1-second bin

3. Calculate the correlation of the participant’s binned time series with the
   group norm

4. Scale the correlation using the minimal and maximal possible correlation
   given the number of segmentation responses


## Conditions

The experiment will be a within-subject design, so all participants will be in
all conditions. Three variables are systematically and orthogonally varied
across the stimuli set each participant gets to see:

1. Trend (metric): -1, -0.5, 0.5, 1 (only applied to the regions of the map
   selected to change)

2. Framing: Invasive vs. Endangered

3. Spatial Pattern: Clustered vs. Distributed

# Hypothesis

This study aims to investigate whether segmentation agreement increases when the
regions of the maps that depict change over time are clustered together rather
then spread out across the map.

Based on the results of a previous study, we expect the marginal effect of trend
on segmentation agreement to be negative if the species is framed as endangered
and positive if it is framed as invasive. We further expect this difference to
be more pronounced for maps in which the changing regions are clustered together
than for those in which they are spread out.

```{r hypothesis, fig.align = "center", fig.height = 5, fig.width = 7}
design <- expand.grid(
  trend       = c(-1, -0.5, 0.5, 1),
  framing     = c(-1, 1),  # 1 = Invasive, -1 = Endangered
  spatialdist = c(0, 1)    # 0 = Distributed, 1 = Clustered,
)

# Intercept
f_intercept <- 0.597

# Main effects
f_trend        <- 0.0
f_spatialdist <- 0.0
f_framing      <- 0.0

# Interactions
f_trend_spatialdist         <- 0.0
f_trend_framing             <- 0.0
f_framing_spatials_dist     <- 0.0
f_trend_spatialdist_framing <- 0.02

design$seg_agree <-
  f_intercept +
  f_trend * design$trend +
  f_spatialdist * design$spatialdist +
  f_framing * design$framing +
  f_trend_spatialdist * design$trend*design$spatialdist +
  f_trend_framing * design$trend * design$framing +
  f_trend_spatialdist_framing * design$trend * design$spatialdist * design$framing

design$framing <- factor(design$framing,
                         levels = c(-1, 1),
                         labels = c("Endangered", "Invasive"))

design$spatialdist <- factor(design$spatialdist,
                             levels = c(0, 1),
                             labels = c("Distributed", "Clustered"))

# Hypothesis plot
lattice::xyplot(seg_agree ~ trend | framing,
                design,
                groups = spatialdist,
                type = "l",
                xlab = "Trend",
                ylab = "Mean Segmentation Agreement",
                auto.key = list(space = "top", column = 2))
```

# Data simulation

```{r datasim}
# Random effect subjects
subject_sd             <- 0.1
subject_trend_sd       <- 0.03
subject_framing_sd     <- 0.01
subject_spatialdist_sd <- 0.01
residual_sd            <- 0.2

n <- 175

dat <- expand.grid(
  subj_id     = c(paste("subj", stringr::str_pad(1:n, 2, "left",0), sep = "_")),
  trend       = c(-1, -0.5, 0.5, 1  ),
  framing     = c(-1, 1),  # 1 = Invasive, -1 = Endangered
  spatialdist = c(0, 1)    # 0 = Distributed, 1 = Clustered,
)
 
subject_intercepts <- data.frame(
  subj_id     = c(paste("subj", stringr::str_pad(1:n, 2, "left",0), sep = "_")),
  subj_interc = rnorm(n, mean = 0, sd = subject_sd)
)
subject_slope_trend <- data.frame(
  subj_id     = c(paste("subj", stringr::str_pad(1:n, 2, "left",0), sep = "_")),
  subj_interc = rnorm(n, mean = 0, sd = subject_trend_sd)
)
subject_slope_framing <- data.frame(
  subj_id     = c(paste("subj", stringr::str_pad(1:n, 2, "left",0), sep = "_")),
  subj_interc = rnorm(n, mean = 0, sd = subject_framing_sd)
)
subject_slope_spatialdist <- data.frame(
  subj_id     = c(paste("subj", stringr::str_pad(1:n, 2, "left",0), sep = "_")),
  subj_interc = rnorm(n, mean = 0, sd = subject_spatialdist_sd)
)

dat <- dat |>
  group_by(subj_id) |>
  mutate("resp" := f_intercept +
                   # main effects
                   f_trend*trend +
                   f_spatialdist * spatialdist +
                   f_framing*framing +
                   # interactions
                   f_trend_spatialdist * trend * spatialdist +
                   f_trend_framing * trend * framing +
                   f_framing_spatials_dist * framing * spatialdist +
                   f_trend_spatialdist_framing * trend * spatialdist * framing +

                   # random effects
                   subject_intercepts[cur_group_id(), "subj_interc"] +
                   subject_slope_trend[cur_group_id(), "subj_interc"] * trend +
                   subject_slope_framing[cur_group_id(), "subj_interc"] * framing +
                   subject_slope_spatialdist[cur_group_id(), "subj_interc"] * spatialdist +
                   rnorm(n(), mean = 0, sd = residual_sd)) |>
  ungroup()
```

# Descriptives

```{r descriptives}
str(dat)
summary(dat)

# Aggregate data for plot
datm <- dat |>
  group_by(framing,  trend, spatialdist) |>
  summarize(mean_agreement = mean(resp))

datm$framing <- factor(datm$framing,
                       levels = c(-1, 1),
                       labels = c("Endangered", "Invasive"))

datm$spatialdist <- factor(datm$spatialdist,
                           levels = c(0, 1),
                           labels = c("Distributed", "Clustered"))

lattice::xyplot(mean_agreement ~ trend | framing,
                datm,
                groups = spatialdist,
                type = c("p", "r"),
                xlab = "Trend",
                ylab = "Mean Segmentation Agreement",
                auto.key = list(space = "top", column = 2))
```

# Analyses

```{r analysis}
m1 <- lmer(resp ~ framing * trend * spatialdist + (trend + framing + spatialdist | subj_id),
           data = dat)

summary(m1)
```

